---
title: Data
author: Sebastian Zeki
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  rmarkdown::html_document:
    theme: lumen
    toc: true
    toc_depth: 5
    css: style.css
vignette: >
  %\VignetteIndexEntry{EndoMineR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
library(pander)
library(EndoMineR)
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

##Overview

It is envisaged that different users will start at different points of their data preparation. THis section is intended to explain the fake data I have created so the type of data used for the examples can be better understood.

There are several data files used. These are detailed below

## Gastroscopy

####  Raw datasets:

#####TheOGDReportFinal 
A dataset containing fake upper GI endoscopy reports. The report field is provided as a whole report without any fields having been already extracted. There are 2000 rows

#####PathDataFrameFinal 
A dataset containing fake upper GI pathology reports. The report field is provided as a whole report without any fields having been already extracted. There are 2000 rows

####Pre-extracted datasets:

#####Myendo 
This has been extracted using the Extractor method as follows from the raw text within Mypath:

```{r MyendoExtract, eval = FALSE}
mywords <- c("OGDReportWhole","HospitalNumber","PatientName",
             "GeneralPractitioner","Dateofprocedure","Endoscopist",
             "Secondendoscopist","Medications","Instrument","ExtentofExam",
             "Indications","ProcedurePerformed","Findings")
Extractor(TheOGDReportFinal,"OGDReportWhole",mywords)
```

####Mypath 

This has been extracted using the Extractor method as follows from the raw text within Mypath:

```{r MypathExtract, eval = FALSE}
mywords<-c("HospitalNumber","PatientName","DOB","GeneralPractitioner",
           "Dateofprocedure","ClinicalDetails","Macroscopicdescription",
           "Histology","Diagnosis")
Extractor(PathDataFrameFinal,"PathReportWhole",mywords)
```

The original dataset has also been added as "PathReportWhole",

##Colonoscopy

###Raw datasets

####ColonFinal

A dataset containing fake lower GI endoscopy reports. The report field is provided as a whole report without any fields having been already extracted. There are 2000 rows

####PathDataFrameFinalColon

A dataset containing fake lower GI pathology reports. The report field is provided as a whole report without any fields having been already extracted. There are 2000 rows.
