---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
#---POSTER SIZE & DEFAULT FONT---#
poster_height: "38in" # height in inches of poster
poster_width: "45in" # width in inches of poster
font_family: "Palatino" # choose from typical html fonts (example: "Palatino")
#---TITLE BOX OPTIONS---#
#ESSENTIALS
title: 'Using posterdown to create a poster about gastroenterological things'
author: 
  - 'S.Zeki^1^'
affiliation: "^1^Department of Gastroenterology, St Thomas' Hospital, London SE1 7EH"

#STYLE & FORMATTING
titlebox_bgcol: "#008080"  #Colour of the Title Box background
titlebox_bordercol: "#0b4545" #Colour of the title Box border.
titlebox_shape: "" #ONLY posterdown_PDF
titlebox_borderwidth: "1cm"
titletext_fontfamily: Helvetica
title_textcol: "#ffffff" #colour of title text
author_textcol: "#0b4545" # Colour of author text
affiliation_textcol: "#FFFFFF" # Colour of affiliation text
title_textsize: "75pt"         # Poster title fontsize
author_textsize: "50pt"       # Author list font size
affiliation_textsize: "40pt"  # Affiliation font size


#ADDING LOGOS
logoleft_name: "Figures/GSTTLogo.PNG"
logoright_name: "Figures/GSTTLogo.PNG"



#---POSTER BODY OPTIONS---#
body_bgcol: "#ffffff" #colour of the poster main background
body_textsize: "35px"    # Size of the main poster body text
body_textcol: "#000000" # Colour of main text in the body of poster
column_numbers: 4 # Number of columns that the poster has
column_margins: "0.5in" # Margin spacing for columns
columnline_col: "#008080" #colour 
columnline_width: "1mm" #width of line between each column
columnline_style: dashed #style of the column line seperator.


#SECTION TITLE STYLING
sectitle_textcol: "#ffffff" # Colour of the poster section titles
sectitle_bgcol: "#0b4545" # Colour of the section title box
sectitle_bordercol: "#0b4545" # Colour of the border around the section title box.
sectitle_borderwidth: "2mm" # Thicknes of the section title box border
sectitle_boxshape: "4mm 0mm" # Changes the shape of the section title box.
sectitle2_textcol: "#0b4545" # Color of 2nd level header text.
#--OTHER--#
link_col: "#0b4545" #colour of other links within the poster
#--Standard Options--#
output: 
  posterdown::posterdown_html:
    self_contained: false
    pandoc_args: --mathjax
    highlight: tango
    number_sections: true
bibliography: MyBib.bib
---

```{r, include=FALSE}
knitr::opts_chunk$set(results = 'asis', echo = FALSE, comment = FALSE,  warning = FALSE, message = FALSE, fig.align = 'center')
library(knitr)
library(here)
library(EndoMineR)
library(DiagrammeR)
library(CodeDepends)
```

<!--
Bellow you will find typical RMarkdown syntax which should fill in the columns for the poster from top to bottom left to right. The more information you have the more will be filled. Please let me know if you have feedback at brentthorne18@gmail.com or at github.com/brentthorne/posterdown.
ENJOY! XD
-->

# Introduction

Patients with Barrett's oesophagus (BE) undergo regular endoscopic surveillance with a view to earlier oesophageal adenocarcinoma detection. Quality monitoring of this programme relies on manual extraction which, given its laboriousness, is a significant hindrance to robust, large scale and reproducible quality monitoring.

EndoMineR, an open source package written in R, has been developed specifically to automate the extraction of data from endoscopic and associated pathology reports^1^.It contains functions to clean, format and extract elements from free text and perform quality metrics for a range of conditions including in BE. 

# Aim 
We decided to assess the accuracy of the BE extraction algorithms within EndoMineR, for both endoscopic and pathological elements of BE using only pathology reports as input. This is the 'worst case scenario' input data. The functions being assessed were: 1. The extraction of a Prague score, 2. The extraction of the worst pathology grade, 3. The site of biopsied tissue, 4. The site and type of any therapy in the upper GI tract.





## Import the data
```{r dataImport,echo=FALSE}
load(file=here("data", "TheOGDReportFinal.rda"))
load(file=here("data", "PathDataFrameFinal.rda"))
```


```{r echo=FALSE}
read_chunk(here("inst","TemplateProject","munge", "PreProcessing.R"))
```


```{r dataClean,echo=FALSE}
```


```{r dataMerge,echo=FALSE}
```


```{r dataClean2,echo=FALSE}
```


```{r dataExtrapolate,echo=FALSE}
```


```{r dataExtrapolate2,echo=FALSE}
```


```{r dataAnalysis_DisSpecific}
```

# Methods 
160 patient episodes between x and ` with full text pathology data only were acquired from x departments in central London as a training set. Validation was performed on a further 100 pathology reports. The therapy algorithm was performed on a further 100 reports.
You can even use the bibliography the same way [@turnerControlsWaterBalance2014].


```{r dataAnalysis_Generic,out.width="100%", cache=FALSE, fig.cap='Metric of one thing versus another', comment=FALSE}
```


## Show the Consort diagram
```{r dataDiagrammeR,out.width="100%", cache=FALSE, fig.cap='Consort diagram of included cases', comment=FALSE}
```

## Show the CodeDepends
```{r codeDepends,out.width="100%", cache=FALSE, fig.cap='Demonstration of how the code is linked together', comment=FALSE}
```


```{r, mytable, out.width='80%'}
library(kableExtra)
knitr::kable(iris[1:5, 1:4], caption = 'Table caption.',align = 'c',"html") %>% 
  kable_styling(full_width = TRUE)
```


```{r, out.width='100%', fig.height=5, fig.cap='Using ggplot and patchwork to generate a layout of multiple plots in one figure. The iris dataset was used to generate (a) a line graph, (b) a scatterplot, and (c) a boxplot all together!'}

```


**Results** Reports were written by x different pathologists. The readability index of all the text, using the Fleisch-Kincaid readability index was xindicating an average grammatical complexity.  The results are displayed in Table 1.Sensitivity was excellent for all algorithms especially given the difficult input text. A reduction in specificity in the detection of worst pathology occurred because of dual reporting of colonoscopy and gastroscopy tissue which also affected the sensitivity of the Pathology Site detection. A variability in how intestinalisation was reported also affected the specicificity.

Usually you want to have a nice table displaying some important results that you have calcualated. In posterdown this is as easy as using the `kable` table formatting you are probably use to as per typical `RMarkdown` formatting. I suggesting checking out the `kableExtra` package and its in depth documentation on customizing these tables found [here](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) [@kableExtra2019]. Hopfully I can make this with an inline refernce like, Table \@ref(tab:mytable).


**Conclusion** 1. Reproducible extraction of BE parameters can be done from semi-structured text. 2. Further improvements using parts of speech tagging and term mapping will improve the results. 3. Such data extraction will allow for upstream automation of quality monitoring and governance and novel metrics in BE as well as other gastroenterological conditions.

**References**
<font size="1"> 1. Zeki S, (2018). EndoMineR for the extraction of endoscopic and associated pathology data from medical reports. Journal of Open Source Software, 3(24), 701 </font>


# References
