---
title: "TemplatePresentation"
author: "Sebastian Zeki"
date: "4/23/2019"
output: ioslides_presentation

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = 'asis', echo = FALSE, comment = FALSE,  warning = FALSE, message = FALSE, fig.align = 'center')
library(knitr)
library(here)
library(EndoMineR)
library(DiagrammeR)
library(CodeDepends)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Import the data
```{r dataImport,echo=FALSE}
load(file=here("data", "TheOGDReportFinal.rda"))
load(file=here("data", "PathDataFrameFinal.rda"))
```


```{r echo=FALSE}
read_chunk(here("inst","TemplateProject","munge", "PreProcessing.R"))
```

## Pre-process both the pathology and the endoscopy text:
```{r dataClean,echo=FALSE}
```

## Merge endoscopy and pathology text
```{r dataMerge,echo=FALSE}
```

## Clean the endoscopist and instrument columns
```{r dataClean2,echo=FALSE}
```

## Extract number of biopsies, biopsy size, biopsy location and medications from the text
```{r dataExtrapolate,echo=FALSE}
```

## Extract endoscopy events
```{r dataExtrapolate2,echo=FALSE}
```

## Extract some generic analyses
```{r dataAnalysis_DisSpecific,echo=FALSE}
```

## Extract some disease specific analyses
```{r dataAnalysis_Generic,echo=FALSE}
```

## Show the Consort diagram
```{r dataDiagrammeR,echo=FALSE}
```

## Show the CodeDepends
```{r codeDepends,echo=FALSE}
```

