---
title: "Template Script"
author: "Sebastian Zeki"
date: "20/11/2018"
output: html_document

---


To create this report within a project start by creating a project with library('ProjectTemplate'). This will create the skeleton project structure. Using 'usethis' or 'rrtools' may also help if you want to create collaborative projects using github as it should set up the github project for you.
you can then drop this markdwon script in the project and use the chunk structure here to run your scripts

### Title and abstract

**$\color{red}{\text{Always use PICO}}$.**
**$\color{red}{\text{1.Patient population}}$.**
**$\color{red}{\text{2.Intervention}}$.**
**$\color{red}{\text{3.Comparison}}$.**
**$\color{red}{\text{4.Outcome}}$.**
**$\color{red}{\text{5. Further}}$.**


In patients being investigated for GORD with a negative impedance study for GORD who go on to have a BRAVO study, what proportion have a positive BRAVO for GORD condition




```{r setup, warning = FALSE, message = FALSE,include=FALSE}
knitr::opts_chunk$set(results = 'asis', echo = FALSE, comment = FALSE,  warning = FALSE, message = FALSE, fig.align = 'center')
library(knitr)
library(here)
library(EndoMineR)
library(DiagrammeR)
library(CodeDepends)
```

### Introduction

**$\color{red}{\text{B.Introductory sentence 1: Define the disease condition}}$.**
**$\color{red}{\text{C.Introductory sentence 2: The overview of the disease condition}}$.**
**$\color{red}{\text{D.Introductory sentence 3: The problem being addressed}}$.**
**$\color{red}{\text{E.Define the Dates}}$.**
**$\color{red}{\text{F.Define the Location}}$.**


Aperistaltic oesophagus is a term that covers all conditions where achalasia is absent and there is no peristaltic activity. Aperistaltic oesophagus remains a difficult condition to treat with no specific and effective treatments. The number of non-achalasia aperistaltic oesophagus patients is unknown.Although often isolated from outpatients in veterinary clinics, there is concern that MRSP follows a veterinary-hospital associated epidemiology

### Aim

The aim of the study was to assess the number of patients with aperistaltic oesophagus without achalasia and ascertain their demographics


### Methods

**$\color{red}{\text{F.Define the Eligibility}}$.**
**$\color{red}{\text{F.Define the Exclusion}}$.**
**$\color{red}{\text{G.Define the Sampling}}$.**
**$\color{red}{\text{H.Define the data cleaning methodology methods}}$.**

The results of high resolution studies performed in patients with cough as part of the presenting complaint to the gastroenterology service at Guy's and St Thomas' NHS Trust were examined retrospectively between June 2008 and January 2018 (Get the dates from the minimum of the VisitDate column `r 1+1 `). All patients selected were adults over the age of 18. Ethics was approved (IRAS number) and by the local ethics board. Patients with incomplete data sets were exluded from the study. Where x was not available fromt the report, the raw file was located and manually analysed. Patients were selected according to Chicago Classification v3. Patients symptoms were selected from self reporting at impedance if this was done at the same time as high resolution manometry. If this was not the case then the symptoms were extracted either from the final report summarising the patient's condition and physiological diagnosis, or from the clinicians details when ordering the investigation. This involves how the relevant columns were accordionised (eg categorising the patients/and how subgroups were formed).



```{r dataImport,echo=FALSE}
load(file=here("data", "TheOGDReportFinal.rda"))
load(file=here("data", "PathDataFrameFinal.rda"))
```


```{r echo=FALSE}
read_chunk(here("inst","TemplateProject","munge", "PreProcessing.R"))
```


```{r dataClean,echo=FALSE}
```


```{r dataMerge,echo=FALSE}
```


```{r dataClean2,echo=FALSE}
```

```{r dataExtrapolate,echo=FALSE}
```


```{r dataExtrapolate2,echo=FALSE}
```


```{r dataAnalysis_DisSpecific,echo=FALSE}
```


```{r dataAnalysis_Generic,echo=FALSE}
```


```{r dataDiagrammeR,echo=FALSE}
```


```{r codeDepends,echo=FALSE}
```

## Results

**$\color{red}{\text{I.Define the Statistical methods}}$.**
**$\color{red}{\text{I.Describe demographics}}$.**
**$\color{red}{\text{J.Then describe level one results}}$.**
**$\color{red}{\text{K.Then describe subset results}}$.**
**Choose your statistical analysis**

The mean age was 57.3 ± 17 years, and 62.3% of the subjects were male. 
The age-adjusted incidence rates were 13.8 (non-lobar) and 4.9 (lobar) per 100,000 person-years.
In the final model, cats (compared to dogs, OR 18.5, 95% CI 1.8–188.0, P = 0.01), animals that had been hospitalised (OR 104.4, 95% CI 21.3–511.6, P < 0.001), or visited veterinary clinics more frequently (>10 visits OR 7.3, 95% CI 1.0–52.6, P = 0.049) and those that had received topical ear medication (OR 5.1, 95% CI 1.8– 14.9, P = 0.003) or glucocorticoids (OR 22.5, 95% CI 7.0–72.6, P < 0.001) were at higher risk of MRSP infection, whereas S. pseudintermedius isolates from ears were more likely to belong to the MSSP group (OR 0.09, 95% CI 0.03–0.34, P < 0.001).


## Discussion

**$\color{red}{\text{L.Sentence 1: To be decided}}$.**
**$\color{red}{\text{M.Sentence 2: To be decided}}$.**
**$\color{red}{\text{N.Sentence 3: To be decided}}$.**

These results indicate an association of MRSP infection with veterinary clinic/hospital
settings and possibly with chronic skin disease.


## Limitations
**$\color{red}{\text{O.Sentence 1: To be decided}}$.**

There was an unexpected lack of association between MRSP and antimicrobial therapy; this requires further investigation .(Lehner et al., 2014).






